---
title: "Brazilian Stock Market Analysis: Econophysics and Tsallis Entropy"
author: "Generated Report"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(quantmod)
library(zoo)
library(ggplot2)
library(dplyr)
library(gridExtra)
```

```{r load-data}
# Source the analysis script
source("analyze_brazilian_stocks.R")
# Run the analysis
results <- analyze_brazilian_stocks()
cor_matrix <- results$cor_matrix
entropy_plot <- results$entropy_plot
crisis_events <- results$crisis_events
boom_events <- results$boom_events
crisis_df <- results$crisis_df
crisis_plot <- results$crisis_plot
ibov_returns <- results$ibov_returns
ibov_close <- results$ibov_close
stock_closes <- results$stock_closes
stocks <- results$stocks
```

# Introduction

This report analyzes the Brazilian stock market using principles of econophysics and Tsallis entropy. It studies correlations between major companies and identifies crisis events in the IBOVESPA index.

# Correlation Analysis

The correlation matrix between selected Brazilian stocks is shown below:

```{r cor-matrix}
print(cor_matrix)
```

## Tsallis Entropy of Correlations

```{r entropy-plot}
entropy_plot
```

# Crisis Events

## Summary

- Number of crisis days detected: `r sum(crisis_events, na.rm = TRUE)`
- Percentage of crisis days: `r round(sum(crisis_events, na.rm = TRUE) / length(crisis_events) * 100, 2)`%

## List of Crisis Dates

The specific dates of crisis events are:

```{r crisis-dates}
crisis_dates <- index(ibov_returns)[crisis_events]
crisis_dates <- crisis_dates[!is.na(crisis_dates)]
print(as.Date(crisis_dates))
```

# Boom Events

## Summary

- Number of boom days detected: `r sum(boom_events, na.rm = TRUE)`
- Percentage of boom days: `r round(sum(boom_events, na.rm = TRUE) / length(boom_events) * 100, 2)`%

## List of Boom Dates

The specific dates of boom events are:

```{r boom-dates}
boom_dates <- index(ibov_returns)[boom_events]
boom_dates <- boom_dates[!is.na(boom_dates)]
print(as.Date(boom_dates))
```

## Crisis Event Plots

For the 5 most severe crisis events (by magnitude of daily return), we show the IBOVESPA index behavior 30 days before and after the event.

```{r top-crises, fig.width=12, fig.height=20}
# Get crisis returns
crisis_returns <- ibov_returns[crisis_events]
# Sort by most negative
top_crisis_indices <- order(crisis_returns)[1:5]
top_crisis_dates <- crisis_dates[top_crisis_indices]
top_crisis_dates <- top_crisis_dates[!is.na(top_crisis_dates)]

# Function to plot around crisis
plot_crisis <- function(crisis_date, days_before = 30, days_after = 30) {
  start_date <- as.Date(crisis_date) - days_before
  end_date <- as.Date(crisis_date) + days_after
  subset_close <- window(ibov_close, start = start_date, end = end_date)
  if (length(subset_close) == 0) return(NULL)
  df <- data.frame(Date = index(subset_close), Close = as.numeric(subset_close))
  ggplot(df, aes(x = Date, y = Close)) +
    geom_line() +
    geom_vline(xintercept = as.numeric(crisis_date), color = "red", linetype = "dashed") +
    labs(title = paste("IBOVESPA around", as.Date(crisis_date)),
         x = "Date", y = "Index Level") +
    theme_minimal()
}

# Generate plots
plots <- lapply(top_crisis_dates, plot_crisis)
# Arrange in grid
do.call(grid.arrange, c(plots, ncol = 1))
```

# Overall Crisis Plot

```{r overall-crisis-plot}
crisis_plot
```

# Boom Event Plots

For the 5 most significant boom events (by magnitude of daily return), we show the IBOVESPA index behavior 30 days before and after the event.

```{r top-booms, fig.width=12, fig.height=20}
# Get boom returns
boom_returns <- ibov_returns[boom_events]
# Sort by most positive
top_boom_indices <- order(boom_returns, decreasing = TRUE)[1:5]
top_boom_dates <- boom_dates[top_boom_indices]
top_boom_dates <- top_boom_dates[!is.na(top_boom_dates)]

# Function to plot around boom
plot_boom <- function(boom_date, days_before = 30, days_after = 30) {
  start_date <- as.Date(boom_date) - days_before
  end_date <- as.Date(boom_date) + days_after
  subset_close <- window(ibov_close, start = start_date, end = end_date)
  if (length(subset_close) == 0) return(NULL)
  df <- data.frame(Date = index(subset_close), Close = as.numeric(subset_close))
  ggplot(df, aes(x = Date, y = Close)) +
    geom_line() +
    geom_vline(xintercept = as.numeric(boom_date), color = "green", linetype = "dashed") +
    labs(title = paste("IBOVESPA around Boom on", as.Date(boom_date)),
         x = "Date", y = "Index Level") +
    theme_minimal()
}

# Generate plots
plots <- lapply(top_boom_dates, plot_boom)
# Arrange in grid
do.call(grid.arrange, c(plots, ncol = 1))
```

# Individual Stock Behavior During Crises

For each of the top 5 crisis events, the following plots show the adjusted closing prices of all selected Brazilian stocks 30 days before and after the crisis. Each facet represents a different stock, allowing comparison of how individual companies were affected.

```{r stock-crises, fig.width=14, fig.height=10}
for (crisis_date in top_crisis_dates) {
  df_list <- lapply(stocks, function(stock) {
    if (is.null(stock_closes[[stock]])) return(NULL)
    start_date <- as.Date(crisis_date) - 30
    end_date <- as.Date(crisis_date) + 30
    subset_close <- window(stock_closes[[stock]], start = start_date, end = end_date)
    if (length(subset_close) == 0) return(NULL)
    data.frame(Date = index(subset_close), Price = as.numeric(subset_close), Stock = stock)
  })
  df_list <- df_list[!sapply(df_list, is.null)]
  if (length(df_list) == 0) next
  df <- do.call(rbind, df_list)
  p <- ggplot(df, aes(x = Date, y = Price)) +
    geom_line() +
    geom_vline(xintercept = as.numeric(crisis_date), color = "red", linetype = "dashed") +
    facet_wrap(~ Stock, scales = "free_y") +
    labs(title = paste("Stock Prices around Crisis on", as.Date(crisis_date)),
         x = "Date", y = "Adjusted Close Price") +
    theme_minimal()
  print(p)
}
```

# Individual Stock Behavior During Booms

For each of the top 5 boom events, the following plots show the adjusted closing prices of all selected Brazilian stocks 30 days before and after the boom. Each facet represents a different stock, allowing comparison of how individual companies performed during positive market events.

```{r stock-booms, fig.width=14, fig.height=10}
for (boom_date in top_boom_dates) {
  df_list <- lapply(stocks, function(stock) {
    if (is.null(stock_closes[[stock]])) return(NULL)
    start_date <- as.Date(boom_date) - 30
    end_date <- as.Date(boom_date) + 30
    subset_close <- window(stock_closes[[stock]], start = start_date, end = end_date)
    if (length(subset_close) == 0) return(NULL)
    data.frame(Date = index(subset_close), Price = as.numeric(subset_close), Stock = stock)
  })
  df_list <- df_list[!sapply(df_list, is.null)]
  if (length(df_list) == 0) next
  df <- do.call(rbind, df_list)
  p <- ggplot(df, aes(x = Date, y = Price)) +
    geom_line() +
    geom_vline(xintercept = as.numeric(boom_date), color = "green", linetype = "dashed") +
    facet_wrap(~ Stock, scales = "free_y") +
    labs(title = paste("Stock Prices around Boom on", as.Date(boom_date)),
         x = "Date", y = "Adjusted Close Price") +
    theme_minimal()
  print(p)
}
```